### 2014-10-22 old_notes/model_07_13.tex
ive added code for tracking mutual information over time and changed figures a bit

The stationary distribution of the estimate will be 
\begin{equation}
X^{(1)}\sim  \label{ddm_stat}
\end{equation}

@@

Let $\lambda_s=m_e- \frac{b}{2}\sqrt{f'\left(\frac{m_f}{m_e}\right)f'\left(\frac{-m_f}{m_e}\right)}$ and $\lambda_{ns}=m_e-  \frac{b}{2}\sqrt{f'\left(\frac{m_f}{m_e}\right)f'\left(\frac{-m_f-b}{m_e}\right)}$ so that $\lambda_s$ is the eigenvalue associated with the equilibrium where Animal 2 signals and $\lambda_{ns}$ is the eigenvalue associated with the equilibrium where Animal 2 is not signaling. 
If we further assume that $f$ is a decreasing sigmoid, then $f'$ is decreasing for $x<-T$ and increasing for $x>-T$ so that $0<\left|f'\left(\frac{-m_f-b}{m_e}\right)\right|<\left|f'\left(\frac{-m_f}{m_e}\right)\right|$.  Therefore, $0>\lambda_s>\lambda_{ns}$.  In general, this suggests that (SMALL!) perturbations around the non-signaling equilibrium, $\left(\frac{-m_f}{m_e},\frac{m_f}{m_e}\right)$, will shrink faster than (SMALL!) perturbations around the signaling equilibrium, $\left(\frac{-m_f-b}{m_e},\frac{m_f}{m_e}\right)$.

@@

We are interested in the effect of signaling feedback, i.e. the $b_s$ parameter, on the dynamics of the model.  We run the algorithm $1000$ times so that we obtain, at each point in time, a distribution of each animal's estimates.  We then compare these distributions when there is and isn't signaling.  In both regimes, each animal's estimates reach a relatively stationary distribution fairly quickly so that we can pick a typical time point at which the distributions are not changing any more.  (A Kolmogorov-Smirnov test for the differences between distributions at different points in time does not allow us to reject the null hypothesis that the distributions are the same, at a significance level of $p=0.05$.)  In Figure (\ref{diff_dists}), we plot the distribution of each animal's estimates at $t=200$ when there is and isn't signaling feedback ($b_{s_1}=b_{s_2}=0$ and $b_{s_1}=b_{s_2}\gneq 0$).  We find that the distributions with and without signaling feedback are quite different.   They are, in fact, statistically distinguishable (K-S test statistic $D=0.68,$ $p<2.2\times 10^{-16}$ for Animal 2 and $D=0.21$, $p<0.05$ for Animal 2).  In both cases, the most noticeable affect of increasing the signaling feedback is increasing the variance of the distributions (from $1.55$ to $81.44$ for Animal 1 and from $1.23$ to $7.02$ for Animal 2).  While the mean for Animal 2 is similar in the two cases ($-10.03$ without feedback and $-9.36$ with feedback), the mean for Animal 1 noticeably increases from $10.04$ to $17.25$.

By comparing distributions of the estimates at a given point in time, we compare the behavior of the stochastic process over many realizations.  We can also compare the behavior of individual trajectories over time.  We hypothesize that by increasing the strength of signaling feedback, we should increase the stability of signaling behavior and it should be less likely that the dominant animal's estimate sink to such an extent that it becomes subordinate.  To test this hypothesis, we count, for each run of the simulation, the number of switches between dominance regimes.  A dominance regime is defined by one animal's estimate being below the signaling threshold $-T$ and the other animal's estimate being above the positive threshold $T$.  If a dominance regime is established with Animal $i$'s estimate above the threshold (dominant) and Animal $i+1$'s estimate below the threshold (subordinate) and subsequently a dominance regime is established with Animal $i$ subordinate and Animal $i+1$ dominant, we count this as one switch.  Fixing the other parameters of the model, for each difference in fighting ability $m_f$ and level of signaling feedback $b_{s_1}=b_{s_2}$, we can find the average number of switches that occur in a run of the model.  This is shown in Figure (\ref{diff_switches}).  As expected, as the difference in fighting ability increases fewer switches occur since Animal 1 is dominant the vast majority of the time.  However, what is less expected is that increasing signaling feedback actually decreases stability.

\begin{figure}
%\begin{subfigure}{.5 \textwidth}
\begin{center}
\includegraphics[width=.85\textwidth]{feedback_estimate_distributions_v1.pdf}
\end{center}
%\caption{}
%\end{subfigure}
\caption{\label{diff_dists} The distribution of estimates for each animal at time $200$, for no signaling feedback and for signaling feedback $b_{s_1}=b_{s_2}=3$.  Other parameters are set as follows: threshold $T=5$, $m_{e_1}=m_{e_2}=-.5$, $n_{e_1}=n_{e_2}=1$, $m_f=5$, $n_f=.5$. Vertical red lines indicate predicted mean for process without feedback, $-m_f/m_{e_1}$ for Animal 1 and $m_f/m_{e_2}$ for Animal 2. }
\end{figure}

\begin{figure}
\begin{subfigure}{1 \textwidth}
\begin{center}
\includegraphics[width=.75\textwidth]{feedback_stability_of_dominance_v1.pdf}
\end{center}
%\caption{}
\end{subfigure}
%\begin{subfigure}{.5 \textwidth}
%\includegraphics[width=1\textwidth]{feedback_stability_of_dominance_v2.pdf}
%\caption{}
%\end{subfigure}
\caption{\label{diff_switches} The number of switches between dominance regimes as a function of the difference in fighting ability, $m_f$, between the animals and the degree to which receiving a signal affects an animal's estimate of its dominance, $b_{s_1}=b_{s_2}$.  Lighter colors correspond to a higher number of switches.  }
\end{figure}

@@

\begin{itemize}
\item More mathematical analysis.
\item Embed the signaling dynamics in a network.  
\begin{itemize}
\item What's the relationship between the distribution of fighting abilities and the distribution of signaling behaviors?  Between fighting abilities and the distribution of power?  And how does that depend on the metric used to calculate power?
\item What's the timescale separation between fighting abilities / fight outcomes / signaling behaviors and the power distribution?  Tradeoff between mutual information between fighting abilities and power distribution and stability of the power distribution?
\end{itemize}
\end{itemize}

>>>>

The stationary distribution of the estimate will be  (I THINK)
\begin{equation}
X^{(1)}\sim N(\exp((m_{e_1}+m_f)t),\frac{n_f^2+n_e^2}{2}t) \label{ddm_stat}
\end{equation}

@@

Let $\lambda_s=m_e- \frac{b}{2}\sqrt{f'\left(\frac{m_f}{m_e}\right)f'\left(\frac{-m_f}{m_e}\right)}$ and $\lambda_{ns}=m_e-  \frac{b}{2}\sqrt{f'\left(\frac{m_f}{m_e}\right)f'\left(\frac{-m_f-b}{m_e}\right)}$ so that $\lambda_s$ is the eigenvalue associated with the equilibrium where Animal 2 signals and $\lambda_{ns}$ is the eigenvalue associated with the equilibrium where Animal 2 is not signaling.  There are two effects of signaling feedbacks on these eigenvalues.  First, when $b\gneq 0$, both $\lambda_s$ and $\lambda_{ns}$ decrease so that the decay to the equilibria is faster when the estimates are influenced by signaling behaviors.

If we further assume that $f$ is a decreasing sigmoid, then $f'$ is decreasing for $x<-T$ and increasing for $x>-T$ so that $0<\left|f'\left(\frac{-m_f-b}{m_e}\right)\right|<\left|f'\left(\frac{-m_f}{m_e}\right)\right|$.  Therefore, $0>\lambda_s>\lambda_{ns}$.  In general, this suggests that (SMALL!) perturbations around the non-signaling equilibrium, $\left(\frac{-m_f}{m_e},\frac{m_f}{m_e}\right)$, will shrink faster than (SMALL!) perturbations around the signaling equilibrium, $\left(\frac{-m_f-b}{m_e},\frac{m_f}{m_e}\right)$.  This effect is greater when the transition from non-signaling to signaling behavior as a function of $x$ is shallower: when the transition is steep, $f'(x)$ is essentially zero for $x>-T$ so that there is only a negligible difference between $f'\left(\frac{-m_f}{m_e},\frac{m_f}{m_e}\right)$ and $f'\left(\frac{-m_f-b}{m_e},\frac{m_f}{m_e}\right)$

@@

We are interested in the effect of signaling feedback, i.e. the $b_s$ parameter, on the dynamics of the model.  We run the algorithm $1000$ times so that we obtain, at each point in time, a distribution of each animal's estimates.  We then compare these distributions when there is and isn't signaling.  In both regimes, each animal's estimates reach a relatively stationary distribution fairly quickly so that we can pick a typical time point at which the distributions are not changing any more.  (A Kolmogorov-Smirnov test for the differences between distributions at different points in time does not allow us to reject the null hypothesis that the distributions are the same, at a significance level of $p=0.05$.)  If we allow the simulation to reach a stationary state, we find that the stationary distributions with and without signaling feedback ($b_{s_1}=b_{s_2}=0$ and $b_{s_1}=b_{s_2}\gneq 0$) are quite different for Animal 1(K-S test statistic, $D=0.995,p<2.2\times10^{-16}$), whereas they are statistically indistinguishable for Animal 2 ($D=0.048, p=0.1995$) (Figure (\ref{diff_dists})).

By comparing distributions of the estimates at a given point in time, we compare the behavior of the stochastic process over many realizations.  We can also compare the behavior of individual trajectories over time.  We hypothesize that by increasing the strength of signaling feedback, we should increase the stability of signaling behavior and it should be less likely that the dominant animal's estimate sink to such an extent that it becomes subordinate.  To test this hypothesis, we count, for each run of the simulation, the number of switches between dominance regimes.  A dominance regime is defined by one animal's estimate being below the signaling threshold $-T$ and the other animal's estimate being above the positive threshold $T$.  If a dominance regime is established with Animal $i$'s estimate above the threshold (dominant) and Animal $i+1$'s estimate below the threshold (subordinate) and subsequently a dominance regime is established with Animal $i$ subordinate and Animal $i+1$ dominant, we count this as one switch.  Fixing the other parameters of the model, for each difference in fighting ability $m_f$ and level of signaling feedback $b_{s_1}=b_{s_2}$, we can find the average number of switches that occur in a run of the model.  As the difference in fighting ability increases fewer switches occur since Animal 1 is dominant the vast majority of the time.  Additionally, the average number of switches between dominance regimes decreases as $b_{s}$ increases, and the effect is greater for when the difference in fighting ability is smaller (Figure \ref{diff_switches}).

\begin{figure}
%\begin{subfigure}{.5 \textwidth}
\begin{center}
\includegraphics[width=.85\textwidth]{feedback_estimate_distributions_v2.pdf}
\end{center}
%\caption{}
%\end{subfigure}
\caption{\label{diff_dists} The distribution of estimates for each animal at time $200$, for no signaling feedback and for signaling feedback $b_{s_1}=b_{s_2}=3$.  Other parameters are set as follows: threshold $T=5$, $m_{e_1}=m_{e_2}=-.5$, $n_{e_1}=n_{e_2}=1$, $m_f=5$, $n_f=.5$. }
\end{figure}

\begin{figure}
\begin{subfigure}{1 \textwidth}
\includegraphics[width=.5\textwidth]{feedback_stability_of_dominance_v2.pdf}
\includegraphics[width=.5\textwidth]{feedback_stability_of_dominance_v5.pdf}
\caption{\label{low_error}}
\end{subfigure}
\begin{subfigure}{1 \textwidth}
\includegraphics[width=.5\textwidth]{feedback_stability_of_dominance_v4.pdf}
\caption{\label{high_error}}
\end{subfigure}
%\begin{subfigure}{.5 \textwidth}
%\includegraphics[width=1\textwidth]{feedback_stability_of_dominance_v2.pdf}
%\caption{}
%\end{subfigure}
\caption{\label{diff_switches} The average number of switches between dominance regimes as a function of the difference in fighting ability, $m_f$, between the animals and the degree to which receiving a signal affects an animal's estimate of its dominance, $b_{s_1}=b_{s_2}$.  In general, even for small differences in fighting ability there are very few switches in dominance regimes.  However, fighting ability can be made small enough to allow switches to occur, in which case the strength of signaling feedbacks affects how frequently they happen.  When the leak rate is relatively low, increases in signaling feedback decrease the number of switches (\ref{low_error}).  When the error rate is high / estimates leak back to $0$ very quickly, increasing the signaling feedback actually increases the number of switches between dominance regimes / decreases the stability of a dominance regime (\ref{high_error}).}
\end{figure}

\section{Linear Learning Formulation}

So far, we have considered how the processes of engaging in fights and sending and receiving signals might affect the estimates two animals make of each other's dominance.  We are interested in understanding these processes on the population level and therefore would like to embed these dynamics in a model of a social network.  In order to make this more analytically tractable, we modify our model into a linear learning model and, for the moment, ignore feedback from signaling into the learning process.  Our main question is how the structure of a social network can facilitate individuals' learning about the states of other nodes in the network.

Suppose each individual has a true fighting ability $a_1,\dots,a_N$ and let $f_{jk}=a_j-a_k$.  Consider an individual in a social network observing its own interactions with other members of the group and the interactions between other members of the group and trying to learn each individual's true fighting ability.  Let $x_j(t)$ be the individual's estimate of individual $j$'s fighting ability at time $t$.  The focal individual will be able to learn about other individual's fighting ability by engaging in fights with individuals it interacts with and observing fights between pairs of individuals both of whom it interacts with.  Let 
$$I_{jk}=\left\{
\begin{array}{cccc}
1 & , & \text{ if the focal individual interacts with both $j$ and $k$}\\
0 & , & \text{ otherwise}
\end{array}\right..
$$
Now, $x_j$ will change according to
\begin{align*}
\dot{x}_j(t)&=\sum_kI_{jk}\left[f_{jk}-(x_j(t)-x_k(t))\right] \text{ for } j\lneq N
\\ \text{ and } x_N(t)&=T-\sum_{j\lneq N}x_j(t),
\end{align*}
where we impose a total on the estimates so that they do not diverge.  Therefore,
\begin{align*}
\dot{x}_j(t)&=\sum_{k\neq N}I_{jk}\left[f_{jk}-(x_j(t)-x_k(t))\right]+I_{jN}\left[f_{jN}-(x_j(t)-x_N(t))\right]
\\&=\sum_{k\neq N}I_{jk}\left[f_{jk}-(x_j(t)-x_k(t))\right]+I_{jN}\left[f_{jN}-(x_j(t)-(T-\sum_{k\neq N}x_k(t)))\right]
\\&=\sum_kI_{jk}f_{jk}-x_j(t)\sum_{k}I_{jk}+\sum_{k\neq N}I_{jk}x_k(t)+TI_{jN}-\sum_{k\neq N}I_{jN}x_k(t)
\\&=\sum_kI_{jk}f_{jk}+TI_{jN}-x_j(t)\sum_kI_{jk}+\sum_{k\neq N}(I_{jk}-I_{jN})x_k(t)
\\&=C_j+\sum_{k}L_{jk}x_k(t) \text{ where } C_j=\sum_{k\neq N}I_{jk}f_{jk}+TI_{jN} \text{ and } 
\end{align*}
$L_{jk}$  is an $n-1\times n-1$ matrix given by
$$
L_{jk}=\left\{
\begin{array}{cccc}
-\sum_lI_{jl}+I_{jj}-I_{jN} & , & \text{ if }j=k
\\ I_{jk}-I_{jN} & , & \text{ if } j\neq k
\end{array}\right..
$$
If $x_k(t)=a_k+\frac{T-\sum_ka_k}{N}$  for all $k$, $\dot{x}(t)=0$.  (If $T=\sum_ka_k$ then $x_k(t)=a_k$ is an equilibrium.)

Stability of the equilibrium is determined by the eigenvalues of $L$.

@@

\begin{itemize}
\item More mathematical analysis.
\begin{itemize}
\item We know that if signaling is constantly present or absent, the stationary distribution of the Ornstein-Uhlenbeck process is just shifted and the variance does not change.  However, we do not know what the likelihood of switching from non-signaling to signaling behavior (or vice versa) and whether the estimates would stay in a non-signaling / signaling regime long enough for a stationary distribution to be established.
\end{itemize}
\item Embed the signaling dynamics in a network.  
\begin{itemize}
\item What's the relationship between the distribution of fighting abilities and the distribution of signaling behaviors?  Between fighting abilities and the distribution of power?  And how does that depend on the metric used to calculate power?
\item What's the timescale separation between fighting abilities / fight outcomes / signaling behaviors and the power distribution?  Tradeoff between mutual information between fighting abilities and power distribution and stability of the power distribution?
\end{itemize}
\end{itemize}

### 2014-10-22 old_notes/model_08_29_12.tex
ive added code for tracking mutual information over time and changed figures a bit

\begin{center}
{\bf \LARGE{Development of a Signaling Network}}
\vspace{10pt}
\\ Eleanor Brush
\\ July 11, 2012
\end{center}

@@

\begin{itemize}
\item More mathematical analysis.
\begin{itemize}
\item We know that if signaling is constantly present or absent, the stationary distribution of the Ornstein-Uhlenbeck process is just shifted and the variance does not change.  However, we do not know what the likelihood of switching from non-signaling to signaling behavior (or vice versa) and whether the estimates would stay in a non-signaling / signaling regime long enough for a stationary distribution to be established.
\end{itemize}
\item Embed the signaling dynamics in a network.  
\begin{itemize}
\item What's the relationship between the distribution of fighting abilities and the distribution of signaling behaviors?  Between fighting abilities and the distribution of power?  And how does that depend on the metric used to calculate power?
\item What's the timescale separation between fighting abilities / fight outcomes / signaling behaviors and the power distribution?  Tradeoff between mutual information between fighting abilities and power distribution and stability of the power distribution?
\end{itemize}
\end{itemize}

>>>>

\begin{center}
{\bf \LARGE{Development of a Signaling Network}}
\vspace{10pt}
\\ Eleanor Brush
\\ August 29, 2012
\end{center}

@@

\begin{itemize}
\item More mathematical analysis.
\begin{itemize}
\item We know that if signaling is constantly present or absent, the stationary distribution of the Ornstein-Uhlenbeck process is just shifted and the variance does not change.  However, we do not know what the likelihood of switching from non-signaling to signaling behavior (or vice versa) and whether the estimates would stay in a non-signaling / signaling regime long enough for a stationary distribution to be established.
\end{itemize}
\item Embed the signaling dynamics in a network.  
\begin{itemize}
\item What's the relationship between the distribution of fighting abilities and the distribution of signaling behaviors?  Between fighting abilities and the distribution of power?  And how does that depend on the metric used to calculate power?
\item What's the timescale separation between fighting abilities / fight outcomes / signaling behaviors and the power distribution?  
	\begin{itemize}
	\item when formulated as a Langevin equation: measure timescales through the expected escape time using martingales
	\item when formulated as a dynamical system: measure timescales through time constants of potential wells.
	\end{itemize}
\item What's the / is there a tradeoff between mutual information between fighting abilities and power distribution and stability of the power distribution?
\end{itemize}
\item Compare to data.
	\begin{itemize}
	\item What are the relative rates of fighting and signals observed / predicted?
	\item Can we explain the data with our model?
	\item Do the monkeys weight fights won and lost equally?  Are they trying to maximize reward, minimize punishment, or are the two weighted equally?
	\item Are the monkeys trying to minimize error tae or decision time? Can we tell if they're being forced to signal?
	\item Is there a sense in which the monkeys are performing optimally? 
	\item (ALSO THEORY QUESTION) How efficiently are the monkeys representing /using the information they're observing about the world?
	\item If we can estimate parameters like the thresholds for signaling, how are they related to other variables in the system?
	\end{itemize}
\end{itemize}
